---
title: "BIMM 143 Class 6 HW Q6 (Git)"
author: "Mi Tran A15556214"
date: "10/26/2021"
output: pdf_document
---
---
title: "BIMM 143 Class 6 HW"
author: "Mi Tran A15556214"
date: "10/25/2021"
output: pdf_document
---
# Analysis of Drug Interactions 

## Given code

```{r}
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

# Question 6: How would you generalize the original code above to work with any set of input protein structures?

## Take s1 for example, the repeated code lines are:
### s1 <- read.pdb("4AKE")  # kinase with drug
### s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
### s1.b <- s1.chainA$atom$b
### plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")

> 1. The inputs to the function are .pdb files, which are being read by the command read.pdb().

> 2. The function I will be creating takes data from the .pbd files and trim it down to what we need. Then it chooses a specific chain and plots the results in a graph.

> 3. The output of the function is a plot for a specified protein, specifically analyzes protein drug interactions by reading any protein PDB data.

## From looking at the repeated code lines, we can combine them into function(), naming this function "pdbplot".

```{r}
# First, we need to load bio3d and read specific PDB files.
library(bio3d)
s1 <- read.pdb("4AKE") #kinase with drug
s2 <- read.pdb("1AKE") #kinase no drug
s3 <- read.pdb("1E4Y") #kinase with drug

# Creating the function "pdbplot"
pdbplot <- function(s1) {
  
  s1.chainA1 <- trim.pdb(s1, chain="A", elety="CA")    # trim the chain
  s1.b <- s1.chainA1$atom$b   # creates new data frame with B-factor values 
  # from the atom data frame
  plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")   # creates a plot of 
  # B-factor by atoms in the chain
  
}
```

## Using the function "pdbplot" that we created to graph s1 

```{r}
pdbplot(s1)
```
```{r}
pdbplot(s2)
```
```{r}
pdbplot(s3)
```


